---
title: "ProjectClean"
author: "Jonathan Lin"
date: "2022-12-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Clean version of the project markdown that we are using

loading the data and packages in
```{r}
# libraries necessary
library(tidyverse)
library(plotly)

#  creating the finalDataset1
finalDataset1 <- draftAndSalaryAndStandings2 %>%
  dplyr::select(Winning.Percentage, Pk, DraftYr, MPG, salary, season, name, team)
row.has.na <- apply(finalDataset1, 1, function(x){any(is.na(x))})
sum(row.has.na)
finalDataset1 <- finalDataset1[!row.has.na,]

# making the MPGcat variable
finalDataset1$MPGcat <- finalDataset1 %>%
  mutate(MPG = cut(MPG,
                    breaks = c(0, 12, 24, 36, Inf),
                    labels = c("Low", "Medium", "High", 
                               "Very High")))

# filtering for the correct draft years
finalDataset1 <- finalDataset1 %>%
  dplyr::filter(season == DraftYr + 3)

# creating the new dataset with + 1 for each draft pick 
newDataset <- finalDataset1
newDataset$Pk <- finalDataset1$Pk + 1
```

defining the function that we will use in our regression (WE ARE NOT USING THIS CHECK LM down BELOW)
```{r}
# square distance function for additive model
sqdist <- function(a, b) {
  (a$Pk - b$Pk)^2/60^2 + (a$MPG - b$MPG)^2/48^2 + (a$Winning.Percentage - b$Winning.Percentage)^2
}
# square distance function for just the pick regression
sqdistPK <- function(a, b) {
  (a$Pk - b$Pk)^2/60^2
}
# predict.local function for weighted additive model that predicts onto a new dataframe
predict.local = function(grid, dataset) {
  mu.hat.local = rep(NA, nrow(grid))
  for(ii in 1:nrow(grid)) { 
    row = grid[ii, ]
    w = exp(-(sqdist(row, dataset)/2))
    model = lm(salary ~ Pk + MPG + Winning.Percentage, weights=w, data=dataset)
    mu.hat.local[ii] = predict(model, newdata=row)
  }
  mu.hat.local
}
# predict function for a regression using just the draft pick with same weights and just the draft pick as the covariate
predict.localDP = function(grid, dataset) {
  mu.hat.local = rep(NA, nrow(grid))
  for(ii in 1:nrow(grid)) { 
    row = grid[ii, ]
    w = exp(-(sqdistPK(row, dataset)/2))
    model = lm(salary ~ Pk + Winning.Percentage + MPG, weights=w, data=dataset)
    mu.hat.local[ii] = predict(model, newdata=row)
  }
  mu.hat.local
}

```

first estimate (only run this if you want the first estimate)
```{r}
# creating the average difference and then graphing it 
ATE2 <- data.frame(lapply(1:60, function(p) {
  grid2 <- expand.grid(Pk = p, Winning.Percentage = 0.75, MPG = 1:48)
  grid3 <- expand.grid(Pk = p, Winning.Percentage = 0.25, MPG = 1:48)
  mean(predict(model, grid2) - predict(model, grid3))
}))

ATEvec2 <- as.vector(ATE2)
AverageDifference <- unlist(ATEvec2)
ggplot() + geom_point(aes(x = 1:60, y = AverageDifference))
```

Creating the fake data
```{r}
# creating the fakedata and generating random salaries for each threshold of winnning percentage vs mpg and pick
fakeData1 <- finalDataset1

fakeData1$PCTCat <- fakeData1$Winning.Percentage

fakeData1 <- fakeData1 %>%
  mutate(PCTCat = cut(PCTCat, 
                      breaks = c(0, 0.5, 1),
                      labels = c("Below", "Above")))

fakeData1$MPGCat <- as.numeric(fakeData1$MPG)

fakeData1 <- fakeData1 %>%
  mutate(MPGCat = cut(MPGCat, 
                   breaks = c(0, 16, 32, 48), 
                   labels = c("Low", "Medium", "High")))

goodhigh <- fakeData1$Winning.Percentage >= 0.5 & fakeData1$MPGCat == "High"
goodmedium <- fakeData1$Winning.Percentage >= 0.5 & fakeData1$MPGCat == "Medium"
goodlow <- fakeData1$Winning.Percentage >= 0.5 & fakeData1$MPGCat == "Low"
badhigh <- fakeData1$Winning.Percentage < 0.5 & fakeData1$MPGCat == "High"
badmedium <- fakeData1$Winning.Percentage < 0.5 & fakeData1$MPGCat == "Medium"
badlow <- fakeData1$Winning.Percentage < 0.5 & fakeData1$MPGCat == "Low"
         
Coverage = data.frame(
  mu = c(1:100),
  bin = c("I")
)
view(Coverage)
epsiloni = rnorm(500)
epsiloni
for (j in 1:100) {
  salaryRandom <- function(x, row) {
  if (x[row,]$MPG <= 16) {
    1000000 * (2 - (1:61)/61)
  } else if (x[row,]$MPG <= 32 & x[row,]$MPG > 16) {
    2000000 * (2 - (1:61)/61)
  } else if (x[row,]$MPG <= 48 & x[row,]$MPG > 32) {
    3000000 * (2 - (1:61)/61)
  }
  }
  range <- function(x, row) {
  if (x[row,]$PCTCat == "Above") {
    2 * (1.1 - (1:61)/61)
  } else if (x[row,]$PCTCat == "Below") {
    1 * (1.1 - (1:61)/61)
  }
  }
  potentialSalary = array(dim = c(500, 61)) 
  for (i in 1:500) {
  mu = salaryRandom(fakeData1, i)
  r = range(fakeData1, i)/2
  potentialSalary[i,] = runif(61, min = mu - r, max = mu + r)
  }
  probabilityPick <- function(x, row) {
  numerator = (1 + x[row,]$MPG/(1:61))
  denominator = sum(numerator)
  numerator / denominator
  }
  for (i in 1:500) {
  probs = probabilityPick(fakeData1, i)
  fakeData1[i,]$Pk = sample(1:61, 1, replace = T, prob = probs)
  }
  fakeData1$salary = 1
  fakeData1$salaryPick1 = 1
  for (i in 1:500) {
  fakeData1[i,]$salary = potentialSalary[i, fakeData1[i,]$Pk]
  fakeData1[i,]$salaryPick1 = potentialSalary[i, min(fakeData1[i,]$Pk + 1, 60)]
  }
  mu = mean(fakeData1$salary - fakeData1$salaryPick1) 
  Coverage[j,]$mu = mu
  sigma.hat = array(0, dim=c(6,6))
  for (i in 1:500) {
  phi2 = phi(fakeData1$MPG, fakeData1$Winning.Percentage, fakeData1$Pk, fakeData1$Pk*fakeData1$Winning.Percentage, fakeData1$Pk*fakeData1$MPG)
  sigma.hat = sigma.hat + phi2 %*% t(phi2)/500
  }
  sigma.hat.inv = solve(sigma.hat)
  epsilon.hat = fakeData1$salary - predict(modelFake, newdata=fakeData1)
  sigma.hat.w = array(0, dim=c(6,6))
  for (i in 1:500) {
  phi2 = phi(fakeData1$MPG, fakeData1$Winning.Percentage, fakeData1$Pk, fakeData1$Pk*fakeData1$Winning.Percentage, fakeData1$Pk*fakeData1$MPG)
  sigma.hat.w = sigma.hat.w + epsilon.hat[i]^2 * phi2 %*% t(phi2)/500
  }
  V = sigma.hat.inv %*% sigma.hat.w %*% sigma.hat.inv
  newFakeData1 <- fakeData1
  newFakeData1$Pk <- newFakeData1$Pk + 1
  u <- rowMeans(phi(fakeData1$MPG, fakeData1$Winning.Percentage, fakeData1$Pk, fakeData1$Pk*fakeData1$Winning.Percentage, fakeData1$Pk*fakeData1$MPG) - phi(newFakeData1$MPG, newFakeData1$Winning.Percentage, newFakeData1$Pk, newFakeData1$Pk*newFakeData1$Winning.Percentage, newFakeData1$Pk*newFakeData1$MPG))
  sigma.hat.u = sqrt(t(u) %*% V %*% u / 500)
  target = target2(muhat)
  if (mu >= target - 2*sigma.hat.u & mu <= target + 2*sigma.hat.u) {
    Coverage[j,]$bin = "Yes"
  } else {
    Coverage[j,]$bin = "No"
  }
}
view(Coverage)
ggplot(data = Coverage) + 
  geom_bar(aes(x = bin)) + 
  labs(x = "Coverage", y = "Frequency")

modelFake = lm(salary ~ Pk + MPG + Winning.Percentage + Pk:Winning.Percentage + Pk:MPG, data=fakeData1)
ATEFake = mean(predict(modelFake, fakeData1) - predict(modelFake, newFakeData1))
ATEFake
coef(modelFake)
betahatFake = c(coef(modelFake))
betahatFake
# int, Pk, MPG, Winning.Percentage, Pk:WP, Pk:MPG

newFakeData1 <- fakeData1
newFakeData1$Pk <- newFakeData1$Pk + 1
```

Confidence interval for regular data
```{r}
# model using continuous variables 
set.seed(1)
model = lm(salary ~ Pk + MPG + Winning.Percentage + Pk:Winning.Percentage + Pk:MPG, data=finalDataset1)
coef(model)

# assigning the coefficients based on model estimation
betahat <- c(705832.737, 19432.653, 80920.621, 414297.286, -40408.833, -1736.515)

# phi function as a function of our features
phi <- function(MPG, Winning.Percentage, Pk, Int1, Int2) {
  rbind(1, Pk, MPG, Winning.Percentage, Pk*Winning.Percentage, Pk*MPG)
}
# int, Pk, MPG, Winning.Percentage, Pk:WP, Pk:MPG

phi1 <- phi(finalDataset1$MPG, finalDataset1$Winning.Percentage, finalDataset1$Pk, finalDataset1$Pk*finalDataset1$Winning.Percentage, finalDataset1$Pk*finalDataset1$MPG)

# mu function as a function of our features
muhat = function(MPG, Winning.Percentage, Pk, Int1, Int2) {
  t(phi(MPG, Winning.Percentage, Pk, Int1, Int2)) %*% betahatFake
}
mu1 <- muhat(finalDataset1$MPG, finalDataset1$Winning.Percentage, finalDataset1$Pk, finalDataset1$Pk*finalDataset1$Winning.Percentage, finalDataset1$Pk*finalDataset1$MPG)
mu2 <- muhat(fakeData1$MPG, fakeData1$Winning.Percentage, fakeData1$Pk, fakeData1$Pk*fakeData1$Winning.Percentage, fakeData1$Pk*fakeData1$MPG)
view(mu1)
view(mu2)

# this is what we're going to use to calculate the target
target <- function(m) {
  mean(m(finalDataset1$MPG, finalDataset1$Winning.Percentage, finalDataset1$Pk, finalDataset1$Pk*finalDataset1$Winning.Percentage, finalDataset1$Pk*finalDataset1$MPG) - m(newDataset$MPG, newDataset$Winning.Percentage, newDataset$Pk, newDataset$Pk*newDataset$Winning.Percentage, newDataset$Pk*newDataset$MPG))
}
target2 <- function(m) {
   mean(m(fakeData1$MPG, fakeData1$Winning.Percentage, fakeData1$Pk, fakeData1$Pk*fakeData1$Winning.Percentage, fakeData1$Pk*fakeData1$MPG) - m(newFakeData1$MPG, newFakeData1$Winning.Percentage, newFakeData1$Pk, newFakeData1$Pk*newFakeData1$Winning.Percentage, newFakeData1$Pk*newFakeData1$MPG))
}
# int, Pk, MPG, Winning.Percentage, Pk:WP, Pk:MPG


target(muhat)
# sigma = function(x) { 1/(1+abs(x)) } dont use this!

# sigma.hat
sigma.hat = array(0, dim=c(6,6))
for (i in 1:500) {
  phi1 = phi(finalDataset1$MPG, finalDataset1$Winning.Percentage, finalDataset1$Pk, finalDataset1$Pk*finalDataset1$Winning.Percentage, finalDataset1$Pk*finalDataset1$MPG)
  sigma.hat = sigma.hat + phi1 %*% t(phi1)/500
}
sigma.hat.inv = solve(sigma.hat)

# sigma.hat.w
epsilon.hat = finalDataset1$salary - predict(model, newdata=finalDataset1) 
sigma.hat.w = array(0, dim=c(6,6))
for (i in 1:500) {
  phi1 = phi(finalDataset1$MPG, finalDataset1$Winning.Percentage, finalDataset1$Pk, finalDataset1$Pk*finalDataset1$Winning.Percentage, finalDataset1$Pk*finalDataset1$MPG)
  sigma.hat.w = sigma.hat.w + epsilon.hat[i]^2 * phi1 %*% t(phi1)/500
}

# V
V = sigma.hat.inv %*% sigma.hat.w %*% sigma.hat.inv
V

# sigma.hat.u
u <- rowMeans(phi(finalDataset1$MPG, finalDataset1$Winning.Percentage, finalDataset1$Pk, finalDataset1$Pk*finalDataset1$Winning.Percentage, finalDataset1$Pk*finalDataset1$MPG) - phi(newDataset$MPG, newDataset$Winning.Percentage, newDataset$Pk, newDataset$Pk*newDataset$Winning.Percentage, newDataset$Pk*newDataset$MPG))

view(u)
sigma.hat.u = sqrt(t(u) %*% V %*% u / 500)
sigma.hat.u

target(mu)

sum(is.na(u))
sum(is.na(V))

# confidence interval 
muhat.CI = target(muhat)
muhat.CI

a = ATE - 2*sigma.hat.u
b = ATE + 2*sigma.hat.u

a
b
# why is this different than our estimate based off of our regression?
```

testing interval for fake data
```{r}
Y = c(1:500)
epsiloni = rnorm(500)
mu <- function(Pk, MPG, WP) {
  Pk + MPG*Winning.Percentage
}

Y = c(1:500)
for (i in 1:500) {
Y[i,] = mu(fakeData1[i,]$Pk, fakeData1[i,]$MPG, fakeData1[i]$Winning.Percentage) + epsilon[i,]
}


```

second estimate (one we are using for draft + 1)
```{r}
# estimate using the original dataset
#ATE <- data.frame(lapply(1, function(p) {
  #mean(predict.localDP(finalDataset1, finalDataset1) - #predict.localDP(newDataset, finalDataset1))
#}))

# new estimate using our interactive model
ATE <- mean(predict(model, finalDataset1) - predict(model, newDataset))
ATE

# estimate using the fakedataset generated with assumptions
#ATEFake <- data.frame(lapply(1, function(p) {
  #mean(predict.localDP(fakeData1, finalDataset1) - #predict.localDP(newFakeData1, finalDataset1))
#}))
ATEFake <- mean(predict(modelFake, fakeData1) - predict(modelFake, newFakeData1))
ATEFake

ATEGraph <- data.frame(lapply(1, function(p) {
  predict(model, finalDataset1) - predict(model, newDataset)
}))


ATEFakeGraph <- data.frame(lapply(1, function(p) {
  predict(modelFake, fakeData1) - predict(modelFake, newFakeData1)
}))

ATEvec <- as.vector(ATEGraph)
ATEvec <- unlist(ATEvec)
ATEvecFake <- as.vector(ATEFakeGraph)
ATEvecFake <- unlist(ATEFakeGraph)

plot1 <- ggplot() + 
  geom_point(aes(x = 1:500, y = ATEvec, color = "red"), show.legend = FALSE) + 
  geom_point(aes(x = 1:500, y = ATEvecFake, color = "blue"), show.legend = FALSE) + 
  labs(x = "Players", y = "Difference in Salary")
  #geom_line(aes(x, y = ATE, color = "red")) +
  #geom_line(aes(x, y = ATEFake, color = "blue"))
plot1
```

Generating more figures
```{r}
plot2 <- ggplot() + 
  geom_point(aes(group = finalDataset1$Pk, x = finalDataset1$Pk, y = finalDataset1$salary)) + 
  labs(color = "Pick", x = "Pick", y = "Salary", title = "Real Data")
plot2

plot3 <- ggplot() +
  geom_point(aes(group = fakeData1$Pk, x = fakeData1$Pk, y = fakeData1$salary, color = fakeData1$Pk)) + 
  labs(color = "Pick", x = "Pick", y = "Salary", title = "Generated Data")
plot3

plot4 <- ggplot() + 
  geom_boxplot(aes(group = finalDataset1$Pk, x = finalDataset1$Pk, y = finalDataset1$salary - fakeData1$salary)) +
  labs(x = "Pick Number", y = "Salary Difference between Real - Generated Data")
plot4

# boxplot of the salary of final data set grouped by pick
plot5 <- ggplot() + 
  geom_boxplot(aes(group = finalDataset1$Pk, x = finalDataset1$Pk, y = finalDataset1$salary)) + 
  labs(x = "Pick", y = "Salary")
plot5

plot6 <- ggplot() + 
  geom_boxplot(aes(group = fakeData1$Pk, x = fakeData1$Pk, y = fakeData1$salary)) +
  labs(x = "Pick", y = "Salary")
plot6

finalDataset1$Predict = ATEvec
finalDataset1$Predict
fakeData1$Predict = ATEvecFake
fakeData1$Predict

# predicted difference between draft pick and salary
plot7 <- ggplot() + 
  geom_boxplot(aes(group = finalDataset1$Pk, x = finalDataset1$Pk, y = finalDataset1$Predict)) + 
  labs(x = "Pick", y = "Difference")
plot7

plot8 <- ggplot() + 
  geom_boxplot(aes(group = fakeData1$Pk, x = fakeData1$Pk, y = fakeData1$Predict)) + 
  labs(x = "Pick", y = "Difference")
plot8

# linear model on top of data
plot9 <- ggplot() + 
  geom_point(aes(group = finalDataset1$Pk, x = finalDataset1$Pk, y = finalDataset1$salary)) + 
  geom_point(color = "red", data = predicted_df, aes(group = finalDataset1$Pk, x = Pick, y = salary_pred)) + 
  labs(x = "Pick", y = "Salary")
plot9

predicted_df <- data.frame(salary_pred = predict(model, finalDataset1), Pick = finalDataset1$Pk)

predicted_df2 <- data.frame(salary_pred = phi1)
```